---
title: "Confirmatory Factor Analysis"
subtitle: "Part I: Single-Factor Models"
author: "Burak Tunca, PhD"
date: Lund, Fall 2018
output: html_document
---

>Note: This document is based on the "semoutput" R package developed by Jason Tsukahara. For more information please see: https://github.com/dr-JT/semoutput

```{r setup, include=FALSE}
#libraries to be loaded
library(knitr)
library(rmarkdown)
library(readr)
library(lavaan)
library(semPlot)
#library(sjPlot)
library(semoutput)
library(resultsoutput)
library(corrplot)
library(PerformanceAnalytics)
library(summarytools)
```

# Confirmatory Factor Analysis

## Dataset

For this exercise we will be using the "Political Democracy" dataset. For more information about this dataset run the following code in the R Console: <?PoliticalDemocracy>

```{r load_data}
#load the political democracy dataset
data <- PoliticalDemocracy

#View first few rows of the dataset
head(data)


#descriptive statistics of the dataset
descriptives(data)
descr(data)
view(dfSummary(data))

#correlations between the variables
round(cor(data),2)

#fancy correlations with the corrplot package
M <- cor(data)
corrplot.mixed(M)

#fancy correlations with the PerformanceAnalytics Package
chart.Correlation(data)
```

## One Factor Models

We would like to model the latent factor "Industrialization in 1960" with the following manifest variables:
  
* x1 The gross national product (GNP) per capita in 1960
* x2 The inanimate energy consumption per capita in 1960
* x3 The percentage of the labor force in industry in 1960
  
```{r CFA for a one factor model}

# factors is a list of the latent factor names. To be used for printing lavaan output to nice looking tables. 
# name the latent factor "ind60"
factors <- c("ind60")

# Specify the model parameters. Name the model "ind60model"
ind60model <- "ind60 =~ x1 + x2 + x3" 

# CFA for our model "ind60model"
ind60fit <- cfa(model=ind60model, data=data, 
                mimic="lavaan", missing="ML", std.lv = FALSE)

summary(ind60fit, standardized=TRUE, fit.measures=TRUE)
```

### Summary Output

```{r one factor CFA Summary}
## This tab will print out summary result tables with the most important information. Fit stats, standardized path values, etc...
sem.sig(ind60fit)
sem.fitmeasures(ind60fit)
sem.factorloadings(ind60fit)
sem.factorcor(ind60fit, factors = factors)
sem.factorvar(ind60fit, factors = factors)
sem.rsquared(ind60fit)
```

### Path Diagram

```{r one factor CFA diagram}
## This tab will print out a model diagram.
semPaths(ind60fit, latents = factors, 
         whatLabels = "std", edge.label.cex = .5, layout = "tree", rotation = 2,
         style = "lisrel", intercepts = FALSE, residuals = TRUE, curve = 1,
         curvature = 3, nCharNodes = 8, sizeMan = 6, sizeMan2 = 3, 
         optimizeLatRes = TRUE, edge.color = "#000000")
```

## Multi-Factor Models

In addition to the "Industrialization in 1960 (ind60)" variable above, we would like to include "Democracy in 1960 (dem60)" and "Democracy in 1965 (dem65)" latent factors using the following manifest variables:

* y1 Expert ratings of the freedom of the press in 1960

* y2 The freedom of political opposition in 1960

* y3 The fairness of elections in 1960

* y4 The effectiveness of the elected legislature in 1960

* y5 Expert ratings of the freedom of the press in 1965

* y6 The freedom of political opposition in 1965

* y7 The fairness of elections in 1965

* y8 The effectiveness of the elected legislature in 1965

```{r CFA for multi-factor models}
# name the latent factors 
factors <- c("ind60", "dem60")

# Specify the model parameters. Name the model "cfamodel"
cfamodel <- "
            ind60 =~ x1 + x2 + x3
            dem60 =~ y1 + y2 + y3 + y4
            "

# CFA for our model 
cfafit <- cfa(model=cfamodel, data=data, 
               mimic="lavaan", missing="ML", std.lv = FALSE)

summary(cfafit, standardized=TRUE, fit.measures=TRUE)
```

### Summary Output

```{r CFA Summary}
## This tab will print out summary result tables with the most important information. Fit stats, standardized path values, etc...
sem.sig(cfafit)
sem.fitmeasures(cfafit)
sem.factorloadings(cfafit)
sem.factorcor(cfafit, factors = factors, standardized = TRUE)
sem.factorvar(cfafit, factors = factors)
sem.rsquared(cfafit)
```

### Path Diagram

```{r CFA diagram}
## This tab will print out a model diagram.
semPaths(cfafit, latents = factors, 
         whatLabels = "std", #view standardized parameters
         edge.label.cex = 0.8, #fontsize 
         style = "lisrel", 
         intercepts = FALSE, residuals = TRUE, curve = 1,
         curvature = 3, nCharNodes = 8, sizeMan = 6, sizeMan2 = 3, 
         optimizeLatRes = TRUE)
```

